% to compile this file into a pdf or ps use the following
% 1) latex whitepaper.tex
% for ps: 2) dvips -K whitepaper -o
% for pdf: 2) ps2pdf whitepaper.ps whitepaper.pdf
\documentclass[12pt]{article}
\usepackage{fancyheadings}
\usepackage{doublespace}
\usepackage{enumerate}
\usepackage{graphics}

% \usepackage{doublespace}

% commands used to control margins
\usepackage{geometry}
\geometry{verbose,letterpaper,lmargin=45mm,rmargin=45mm}
% \textheight=9.3in
\topmargin=.2in

% command used to control between-line spacing
\renewcommand{\baselinestretch}{1}

\pagestyle{fancy}
\cfoot{}
\lhead{}
\rhead{\rm\thepage\ of \pageref{'LastPage'}}
\chead{\emph{Evaluating QoS of Distributed Video Synchrony}}

\title{Process Migration Policy Manager}

\author{
 \textbf{Dan B. Phung, Peppo Valetto, Gail Kaiser}\\
Computer Science Department\\
Columbia University\\
\{phung, valetto, kaiser\}@cs.columbia.edu
}

\date{
\parbox[b][0ex]{0em}{\hspace*{-12.5em}\raisebox{37ex}{\fbox{For submission to 
\emph{ACM-MM 2004}, due 12:00 AM EDT: April 05, 2004.}}}}

keywords: collaboration, mutlimedia synchrony, 

Three approaches: 
- introduce a new and useful tool for synchronized multimedia.  
  : I doubt this is useful


  The increasing popularity of online courses has highlighted the
  problemmatic differences in computing resources of the participating
  student population.  In courses that require collaboration among
  students, these differences in computing resources....
    coursework there is a need
- introduce and discuss the scoring system
  : maybe

- comparison of systems.  
  : will suhit ever finish?

\begin{document}

We only focus on network bandwidth as the commodity resource.
(is this a valid assumption?)

geographically dispersed, varying network resources.

scope: education videos, scoring system.

devils advocate:
- share desktops
  - just use vnc, and have everybody connect
  - just use remote desktop
  - use WebEx




\begin{singlespace}
\maketitle
\end{singlespace}

\begin{abstract}\noindent

keywords: 

\end{abstract}

\section{Introduction} \label{sec:intro}

Use case scenarios.

%% need to find out: 
%% - what's state information is available to the our framework?
%%   > when does a process need to be migrated (when does a proc die?)
%%   > what resources are available (what system can we move to?)

\subsection{Motivation} 

\subsection{Contributions} \label{sec:contrib}
goals: 

\subsection{Evaluation} \label{sec:eval}

\subsection{Quality of Service} \label{sec:qos}

\section{Workflow} \label{sec:wf}

\section{Results} \label{sec:results}
\section{Related Work} \label{sec:related}
\section{Future Work} \label{sec:future}
\section{Conclusion} \label{sec:con}
\section{Acknowledgements} \label{sec:ack}

\label{'LastPage'}
\end{document}
